A general and efficient representation of Ancestral Recombination Graphs

### 2022-09-29 paper.tex
Remove explicit listing of redundant nodes
And move to the caption, so that people don't have to refer back to the figure too much.

% The reader may arrive at the end of this section going, er, so what?
% Try to tackle this head on by getting to the heart of the distinction.
The difference between eARGs and gARGs are subtle, and whether
we choose to interpret nodes as events or genomes
might seem of little practical importance.
However, as we show in the remainder of this article,
there are substantial practical gains
from using the gARG encoding, allowing differential
levels of information about the recombination process to be
represented in a computionally efficient way.
The key to this is not so much in how we interpret nodes
(see Appendix XXX), but specifically in how
we represent the process of genetic inheritance.
The critical distinction is that
the gARG encoding allows us to directly describe
genetic inheritance in multiple non-contiguous genome intervals
between an ancestor and a descendant, something that
cannot be done by storing recombination breakpoints.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{illustrations/ancestry-resolution}
\caption{\label{fig-ancestry-resolution}
The \citet[][Fig.~1]{wiuf1999recombination} example graph as an eARG and
two gARGs with alternative edge annotation schemes.
(A) A conventional Griffiths eARG. Square nodes represent events, with
each recombination (red) annotated with the position of a single breakpoint
along the 7-unit-long chromosome.
(B) A gARG with direct-inheritance edge annotations. Oval nodes represent
genomes and annotations reflect ``local'' transmission of all genetic material
between parent and child. Edges that do not span the full genome are visually
highlighted, showing how breakpoint positions can be encoded.
(C) A gARG with sample-resolved edge annotations. The Hudson-like simplification
algorithm has been used to shorten edge spans to
reflect only the upwards transmission of genetic material from the samples
\textsf{a}, \textsf{b}, and \textsf{c}; very few edges then span the entire genome.
The depiction of the genome within each
node is shaded as in Fig.~\ref{fig-arg-in-pedigree}A to show coalescing
regions of genome back to the MRCA nodes \textsf{k} (regions 0 to 2) and \textsf{p}
(regions 2 to 7). Line widths reflect the genomic span of each edge, highlighting
the relative importance of different transmission pathways to the samples.
}
\end{figure}

@@

It is worth noting that nodes \textsf{g},  \textsf{j}, \textsf{o} and \textsf{q}
in Fig~\ref{fig-ancestry-resolution}C are redundant in that they can
be removed without reducing our knowledge of the genetic
ancestry of the samples. Removing nodes,
both redundant ones and others, is part of a process we call
``simplification''.

>>>>

% The reader may arrive at the end of this section going, er, so what?
% Try to tackle this head on by getting to the heart of the distinction.
The difference between eARGs and gARGs are subtle, and whether
we choose to interpret nodes as events or genomes
might seem of little practical importance.
However, as we show in the remainder of this article,
there are substantial practical gains
from using the gARG encoding, allowing differential
levels of information about the recombination process to be
represented in a computationally efficient way.
The key to this is not so much in how we interpret nodes
(see Appendix XXX), but specifically in how
we represent the process of genetic inheritance.
The critical distinction is that
the gARG encoding allows us to directly describe
genetic inheritance in multiple non-contiguous genome intervals
between an ancestor and a descendant, something that
cannot be done by storing recombination breakpoints.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{illustrations/ancestry-resolution}
\caption{\label{fig-ancestry-resolution}
The \citet[][Fig.~1]{wiuf1999recombination} example graph as an eARG and
two gARGs with alternative edge annotation schemes.
(A) A conventional Griffiths eARG. Square nodes represent events, with
each recombination (red) annotated with the position of a single breakpoint
along the 7-unit-long chromosome.
(B) A gARG with direct-inheritance edge annotations. Oval nodes represent
genomes and annotations reflect ``local'' transmission of all genetic material
between parent and child. Edges that do not span the full genome are visually
highlighted, showing how breakpoint positions can be encoded.
(C) A gARG with sample-resolved edge annotations. The Hudson-like simplification
algorithm has been used to shorten edge spans to
reflect only the upwards transmission of genetic material from the samples
\textsf{a}, \textsf{b}, and \textsf{c}; very few edges then span the entire genome.
The depiction of the genome within each
node is shaded as in Fig.~\ref{fig-arg-in-pedigree}A to show coalescing
regions of genome back to the MRCA nodes \textsf{k} (regions 0 to 2) and \textsf{p}
(regions 2 to 7). Line widths reflect the genomic span of each edge, highlighting
the relative importance of different transmission pathways to the samples. Nodes
\textsf{g},  \textsf{j}, \textsf{o} and \textsf{q} are redundant in that they correspond
to events which have no impact on the genetic ancestry of the samples.
}
\end{figure}

@@

It is worth noting the process of sample resolution can reveal that certain ARG
nodes are \emph{redundant}: if they are removed and the edges adjusted to
pas directly from the node's child to the node's parent, it does not reduce our
knowledge of the samples' genetic ancestry. Removing nodes,
both redundant ones and others, is part of a process we call
``simplification''.
